\LARGE \bf Steering a Swarm Using Global Inputs\\ and Swarm Statistics

### 2017-10-03 journalSwarmControl/realExperiment.tex
preparing to read through the paper!

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Object manipulation with hardware robots}\label{sec:realExperiment}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Our experiments use centimeter-scale hardware systems called \emph{kilobots}.  While those are far larger than the micro scale devices we model, using kilobots allows us to emulate a variety of dynamics, while enabling a high degree of control over robot function, the environment, and data collection. The kilobot, reported by Rubenstein \cite{Rubenstein2012,rubenstein2014programmable}, is a low-cost robot designed for testing collective algorithms with large numbers of robots. It is available as an open-source platform or commercially from K-team  \cite{K-Team2015}.  Each robot is approximately 3 cm in diameter, 3 cm tall, and uses two vibration motors to move on a flat surface at speeds up to 1 cm/s.  Each robot has one ambient light sensor that is used to implement \emph{phototaxis},  moving towards a light source.

@@

The kilobots performed five successful runs manipulating a hexagonal object through an obstacle maze. Videos of these runs are in Extension 2. These hardware experiments represent the results of over 100 hours of trials.
Each trial used 100 kilobots. Trials two through five were performed in a row with no failures in between.  For each trial, fully charged kilobots were placed in the lower left-hand of the workspace, as shown in Fig.~\ref{fig:expSnapShot}.  The moveable object was placed in the lower center of the workspace.  {\sc Matlab} code for vision processing, the policy iteration of \S \ref{subsec:objectpolicy} and the algorithm of \S \ref{sec:AlgObjectManipulation} is available on {\sc Matlab} Central at \cite{Shahrokhi2015MDP}.
Trials were run until the object COM entered the goal region.  The trials ran for \{1465, 3457, 3000, 2162, 2707\} s.  This is $2558\pm771$ s (mean$\pm$std).

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Object manipulation with hardware robots}\label{sec:realExperiment}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Our experiments use centimeter-scale hardware systems called \emph{kilobots}.  While those are far larger than the micro scale devices we model, using kilobots allows us to emulate a variety of dynamics, while enabling a high degree of control over robot function, the environment, and data collection. The kilobot, reported by Rubenstein \cite{Rubenstein2012,rubenstein2014programmable}, is a nonholonomic, low-cost robot designed for testing collective algorithms with large numbers of robots. It is available as an open-source platform or commercially from K-team  \cite{K-Team2015}.  Each robot is approximately 3 cm in diameter, 3 cm tall, and uses two vibration motors to move on a flat surface at speeds up to 1 cm/s.  Each robot has one ambient light sensor that is used to implement \emph{phototaxis},  moving towards a light source.

@@

Even though kilobots are nonholonomic, they performed five successful runs manipulating a hexagonal object through an obstacle maze. Videos of these runs are in Extension 2. These hardware experiments represent the results of over 100 hours of trials.
Each trial used 100 kilobots. Trials two through five were performed in a row with no failures in between.  For each trial, fully charged kilobots were placed in the lower left-hand of the workspace, as shown in Fig.~\ref{fig:expSnapShot}.  The moveable object was placed in the lower center of the workspace.  {\sc Matlab} code for vision processing, the policy iteration of \S \ref{subsec:objectpolicy} and the algorithm of \S \ref{sec:AlgObjectManipulation} is available on {\sc Matlab} Central at \cite{Shahrokhi2015MDP}.
Trials were run until the object COM entered the goal region.  The trials ran for \{1465, 3457, 3000, 2162, 2707\} s.  This is $2558\pm771$ s (mean$\pm$std).

### 2017-10-03 journalSwarmControl/theory.tex
preparing to read through the paper!

We represent particles as holonomic robots that move in the 2D plane. We want to control position and velocity of the particles. 
First, assume a noiseless system containing one robot with mass $m$.
 Our inputs are global forces $[u_x,u_y]$. We define our state vector $\mathbf{x}(t)$ as the $x$ position, $x$ velocity, $y$ position and $y$ velocity.
The state-space representation in standard form is: 
\begin{align}\label{eq:stdform}
\dot{\mathbf{x}}(t)  &=  A \mathbf{x}(t) + B \mathbf{u}(t).
\end{align}

@@

We want to find the number of states that we can control, which is given by the rank of the \emph{controllability matrix}
\begin{align}
\mathcal{C} &= [ B, AB, A^2B, ... , A^{n-1}B ].\\
\textrm{Here }
\mathcal{C}&=\left[
\begin{matrix} 
0 & 0\\
\frac{1}{m} & 0 \\
0 & 0 \\
0 & \frac{1}{m}
\end{matrix}
\,\middle\vert\,
\begin{matrix} 
\frac{1}{m}& 0\\
0 & 0\\
0 & \frac{1}{m}\\
0 & 0
\end{matrix}
\,\middle\vert\,
\begin{matrix} 
0 & 0\\
0 & 0 \\
0 & 0 \\
0 & 0
\end{matrix}
\,\middle\vert\,
\begin{matrix} 
0 & 0\\
0 & 0\\
0 & 0\\
0 & 0
\end{matrix}
 \right],\\
 \quad \textrm{rank}(\mathcal{C})&=4,
\end{align}
and thus all four states are controllable. This section starts by proving independent position control of many robots is not possible, but the mean position can be controlled. We then provide conditions under which the variance of many robots is also controllable.

\subsection{Independent control of many particles is impossible}
In this model, a single particle is fully controllable. For holonomic robots, movement in the $x$ and $y$ coordinates are independent, so for notational convenience without loss of generality we will focus only on movement in the $x$ axis. Given $n$ particles to be  controlled in the $x$ axis, there are $2n$ states: $n$ positions and $n$ velocities. Without loss of generality, assume $m=1$.
Our state-space representation is:
\begin{equation}
\begin{bmatrix}
\dot{x}_{1}\\ 
\ddot{x}_{1}\\
\vdots\\
\dot{x}_{n}\\
\ddot{x}_{n}

\end{bmatrix} = \begin{bmatrix}
0 & 1 & \ldots & 0 & 0 \\
0 & 0 & \ldots& 0 & 0 \\
\vdots &  \vdots & \ddots & \vdots & \vdots \\
0 & 0  & \ldots & 0 & 1 \\
0 & 0 & \ldots& 0 & 0 
\end{bmatrix}  \begin{bmatrix}
x_{1}\\
\dot{x}_{1}\\
\vdots \\
x_{n}\\
\dot{x}_{n}
\end{bmatrix} + \begin{bmatrix}
0\\
1\\
\vdots\\
0\\
1
\end{bmatrix} u_x.
\end{equation}
 However, just as with one particle, we can only control two states because the controllability matrix $\mathcal{C}_n$ has rank two:
\begin{equation}
\mathcal{C}_n=\left[ \begin{matrix} 
0\\
1\\
\vdots\\
0\\
1
\end{matrix}
\,\middle\vert\,
  \begin{matrix} 
1\\
0\\
\vdots\\
1\\
0
\end{matrix}
\,\middle\vert\,
\begin{matrix} 
0\\
0\\
\vdots\\
0\\
0
\end{matrix}
\,\middle\vert\,
\begin{matrix} 
0\\
0\\
\vdots\\
0\\
0
\end{matrix}\,\middle\vert\,
\ldots \right], \quad \textrm{rank}(\mathcal{C}_n)=2.
\end{equation}  
\subsection{Controlling the mean position}\label{sec:controlMeanPosition}
This means \emph{any} number of particles controlled by a global command have just two controllable states in each axis. We cannot arbitrarily control the position and velocity of two or more robots, but have options on which states to control.  %One option is to control the position and velocity of the $j^{th}$ robot. To find a potentially more useful option, 
We create the following reduced order system that represents the mean $x$ position and velocity of the $n$ particles:
\begin{align}
\begin{bmatrix}\nonumber
\dot{\bar{x}} \\
\ddot{\bar{x}}
\end{bmatrix} &= \frac{1}{n} \begin{bmatrix}
0& 1& \ldots &0& 1 \\
0& 0& \ldots &0& 0
\end{bmatrix}
\begin{bmatrix}
x_{1}\\
\dot{x}_{1}\\
\vdots\\
x_{n}\\
\dot{x}_{n}
\end{bmatrix} \\
&+ \frac{1}{n}\begin{bmatrix}
0& 0&  \ldots &0& 0 \\
0& 1&  \ldots &0& 1
\end{bmatrix}\begin{bmatrix} 
0\\
1\\
\vdots\\
0\\
1
\end{bmatrix} u_x.
\end{align}
Thus:
\begin{equation}
\begin{bmatrix}
\dot{\bar{x}} \\
\ddot{\bar{x}}
\end{bmatrix} = \begin{bmatrix}
0& 1 \\
0& 0
\end{bmatrix}
\begin{bmatrix}
\bar{x}\\
\dot{\bar{x}}
\end{bmatrix} + \begin{bmatrix} 
0\\
1
\end{bmatrix} u_x.
\end{equation}

@@

We control mean position with a PD controller that uses the mean position and mean velocity. $[u_x,u_y]^\top$ is the global force applied to each robot:
\begin{align}
u_x &= K_{p}(x_{\rm{goal}} - \bar{x}) + K_{d}(0-\dot{\bar{x}}), \nonumber\\
u_y &= K_{p}(y_{\rm{goal}}  - \bar{y}) + K_{d}(0-\dot{\bar{y}}).  \label{eq:PDcontrolPosition}
\end{align}
 $K_{p}$ is the proportional gain, and $K_{d}$ is the derivative gain.

@@

The variance, $\sigma_x^2,\sigma_y^2$, of $n$ robots' position is computed as:
\begin{align}\label{eq:meanVar}
 \overline{x}(\mathbf{x}) = \frac{1}{n} \sum_{i=1}^n x_{i}, \qquad  %\nonumber \\ 
\sigma_x^2(\mathbf{x}) &= \frac{1}{n} \sum_{i=1}^n (x_{i} - \overline{x})^2,  \nonumber \\ 
 \overline{y}(\mathbf{x}) = \frac{1}{n} \sum_{i=1}^n y_{i}, \qquad  %\nonumber \\ 
\sigma_y^2(\mathbf{x}) &= \frac{1}{n} \sum_{i=1}^n (y_{i} - \overline{y})^2.  
%NOT 1/(N-1) because we are measuring the actual Variance, not the variance of samples drawn from a distribution
\end{align}

@@

If robots with radius $r$ are in a bounded environment with sides of length $[\ell_x, \ell_y]$, the unforced variance asymptotically grows to the variance of a uniform distribution,
\begin{align}
[\sigma_x^2,\sigma_y^2] = \frac{1}{12}[ (\ell_x - 2 r)^2,(\ell_y - 2 r)^2].\label{eq:VarianceUniformDistribution}
\end{align}

 A flat obstacle can be used to decrease variance. Pushing a group of dispersed robots against a flat obstacle will decrease their variance until the minimum-variance (maximum density) packing  is reached. For large $n$, Graham ~\cite{graham1990penny} showed that the minimum-variance packing  for $n$ circles with radius $r$ is 
 \begin{align} \label{eq:optimalvar}
 \sigma^2_{\rm{optimal}}(n,r) \approx   \frac{\sqrt{3}}{\pi} n r^2\approx 0.55 n r^2.
 \end{align} 
% Sloan minimized second moment U = 1/d^2 \sum_{i=1}^{n} || P_i - P ||^2 = (\sqrt{3} n^2)/(4 \pi), where d = 2r.
% Therefore the variance is d^2 (\sqrt{3} n^2)/(4 \pi), or 4^r^2 (\sqrt{3} n^2)/(4 \pi) = r^2 (\sqrt{3} n^2)/(\pi)If we only need to reduce variance in a single axis, the variance can be reduced to zero, given sufficient space.

@@

Similar to the PD controller in Eq.~\eqref{eq:PDcontrolPosition} that controls the mean robot position, a controller to regulate the variance to $\sigma^2_{\rm{ref}}$ is:
\begin{align}
u_x &= K_{p}(x_{\rm{goal}}(\sigma^2_{\rm{ref}}) - \bar{x}) - K_{d}\bar{v}_x + K_{i}(\sigma^2_{\rm{ref}}-\sigma^2_{x}),   \label{eq:PDcontrolVariance}\\
u_y &= K_{p}(y_{\rm{goal}}(\sigma^2_{\rm{ref}})  - \bar{y}) - K_{d}\bar{v}_y + K_{i}(\sigma^2_{\rm{ref}}-\sigma^2_{y}).  \label{eq:PDcontrolVarianceIny}
\end{align}
We call the gain scaling the variance error $K_i$ because the variance, if unregulated, integrates over time. This controller requires a vertical and a horizontal wall.
Eq.~\eqref{eq:PDcontrolVariance} assumes the nearest wall is to the left of the robot at $x=0$, and chooses a reference goal position that in steady-state would have the correct variance according to \eqref{eq:VarianceUniformDistribution}:
\begin{align}
x_{\rm{goal}}(\sigma^2_{\rm{ref}}) = \ell_x/2 = r + \sqrt{3\sigma^2_{\rm{ref}}}.
\end{align}
 If a wall to the right is closer, the signs of $[K_p,K_i]$ are inverted, and the location $x_{\rm{goal}}$ is translated.  A similar argument applies to~\eqref{eq:PDcontrolVarianceIny}.

@@

A key challenge is to select proper values for $\sigma_{\rm{min}}^2$ and $\sigma_{\rm{max}}^2$.  The optimal packing variance was given in \eqref{eq:optimalvar}.
The random packings generated by pushing our robots into corners are suboptimal, so we choose the conservative values: 
%shown in Fig.~\ref{fig:hysteresis}:
\begin{align} \label{eq:sigMaxMin}
 \sigma^2_{\rm{min}} &= 2.5r+ \sigma^2_{\rm{optimal}}(n,r), \nonumber\\
  \sigma^2_{\rm{max}} &= 15r+ \sigma^2_{\rm{optimal}}(n,r).
  \end{align}
%\sigma \approx 0.371258 n r,   \sigma_{min} = n*r*1/2, \sigma_{min} = n*r*3/2, 
%  this result was verified in http://www-math.mit.edu/~tchow/penny.pdf

>>>>

We represent particles as holonomic robots that move in the 2D plane. We want to control position and velocity of the particles. 
First, assume a noiseless system containing one particle with mass $m$.
 Our inputs are global forces $[u_x,u_y]$. We define our state vector $\mathbf{x}(t)$ as the $x$ position, $x$ velocity, $y$ position and $y$ velocity.
The state-space representation in standard form is: 
\begin{align}\label{eq:stdform}
\dot{\mathbf{x}}(t)  &=  A \mathbf{x}(t) + B \mathbf{u}(t).
\end{align}

@@

We want to find the number of states that we can control, which is given by the rank of the \emph{controllability matrix}
\begin{align}
\mathcal{C} &= [ B, AB, A^2B, A^{3}B ].\\
\textrm{Here }
\mathcal{C}&=\left[
\begin{matrix} 
0 & 0\\
\frac{1}{m} & 0 \\
0 & 0 \\
0 & \frac{1}{m}
\end{matrix}
\,\middle\vert\,
\begin{matrix} 
\frac{1}{m}& 0\\
0 & 0\\
0 & \frac{1}{m}\\
0 & 0
\end{matrix}
\,\middle\vert\,
\begin{matrix} 
0 & 0\\
0 & 0 \\
0 & 0 \\
0 & 0
\end{matrix}
\,\middle\vert\,
\begin{matrix} 
0 & 0\\
0 & 0\\
0 & 0\\
0 & 0
\end{matrix}
 \right],\\
 \quad \textrm{rank}(\mathcal{C})&=4,
\end{align}
and thus all four states are controllable. This section starts by proving independent position control of many particles is not possible, but the mean position can be controlled. We then provide conditions under which the variance of many particles is also controllable.

\subsection{Independent control of many particles is impossible}
In this model, a single particle is fully controllable. For holonomic particles, movement in the $x$ and $y$ coordinates are independent, so for notational convenience without loss of generality we will focus only on movement in the $x$ axis. Given $n$ particles to be  controlled in the $x$ axis, there are $2n$ states: $n$ positions and $n$ velocities. Without loss of generality, assume $m=1$.
Our state-space representation is:
\begin{equation}
\begin{bmatrix}
\dot{x}_{1}\\ 
\ddot{x}_{1}\\
\vdots\\
\dot{x}_{n}\\
\ddot{x}_{n}

\end{bmatrix} = \begin{bmatrix}
0 & 1 & \ldots & 0 & 0 \\
0 & 0 & \ldots& 0 & 0 \\
\vdots &  \vdots & \ddots & \vdots & \vdots \\
0 & 0  & \ldots & 0 & 1 \\
0 & 0 & \ldots& 0 & 0 
\end{bmatrix}  \begin{bmatrix}
x_{1}\\
\dot{x}_{1}\\
\vdots \\
x_{n}\\
\dot{x}_{n}
\end{bmatrix} + \begin{bmatrix}
0\\
1\\
\vdots\\
0\\
1
\end{bmatrix} u_x.
\end{equation}
 However, just as with one particle, we can only control two states because the controllability matrix $\mathcal{C}_n$ has rank two:
\begin{equation}
\mathcal{C}_n=\left[ \begin{matrix} 
0\\
1\\
\vdots\\
0\\
1
\end{matrix}
\,\middle\vert\,
  \begin{matrix} 
1\\
0\\
\vdots\\
1\\
0
\end{matrix}
\,\middle\vert\,
\begin{matrix} 
0\\
0\\
\vdots\\
0\\
0
\end{matrix}
\,\middle\vert\,
\begin{matrix} 
0\\
0\\
\vdots\\
0\\
0
\end{matrix}\,\middle\vert\,
\ldots \right], \quad \textrm{rank}(\mathcal{C}_n)=2.
\end{equation}  
\subsection{Controlling the mean position}\label{sec:controlMeanPosition}
This means \emph{any} number of particles controlled by a global command have just two controllable states in each axis. We cannot arbitrarily control the position and velocity of two or more particles, but have options on which states to control.  %One option is to control the position and velocity of the $j^{th}$ robot. To find a potentially more useful option, 
We create the following reduced order system that represents the mean $x$ position and velocity of the $n$ particles:
\begin{align}
\begin{bmatrix}\nonumber
\dot{\bar{x}} \\
\ddot{\bar{x}}
\end{bmatrix} &= \frac{1}{n} \begin{bmatrix}
0& 1& \ldots &0& 1 \\
0& 0& \ldots &0& 0
\end{bmatrix}
\begin{bmatrix}
x_{1}\\
\dot{x}_{1}\\
\vdots\\
x_{n}\\
\dot{x}_{n}
\end{bmatrix} \\
&+ \frac{1}{n}\begin{bmatrix}
0& 0&  \ldots &0& 0 \\
0& 1&  \ldots &0& 1
\end{bmatrix}\begin{bmatrix} 
0\\
1\\
\vdots\\
0\\
1
\end{bmatrix} u_x.
\end{align}
Thus:
\begin{equation}
\begin{bmatrix}
\dot{\bar{x}} \\
\ddot{\bar{x}}
\end{bmatrix} = \begin{bmatrix}
0& 1 \\
0& 0
\end{bmatrix}
\begin{bmatrix}
\bar{x}\\
\dot{\bar{x}}
\end{bmatrix} + \begin{bmatrix} 
0\\
1
\end{bmatrix} u_x.
\end{equation}

@@

We control mean position with a PD controller that uses the mean position and mean velocity. $[u_x,u_y]^\top$ is the global force applied to each particle:
\begin{align}
u_x &= K_{p}(x_{\rm{goal}} - \bar{x}) + K_{d}(0-\dot{\bar{x}}), \nonumber\\
u_y &= K_{p}(y_{\rm{goal}}  - \bar{y}) + K_{d}(0-\dot{\bar{y}}).  \label{eq:PDcontrolPosition}
\end{align}
 $K_{p}$ is the proportional gain, and $K_{d}$ is the derivative gain.

@@

The variance, $\sigma_x^2,\sigma_y^2$, of $n$ particles' position is:
\begin{align}\label{eq:meanVar}
 \overline{x}(\mathbf{x}) = \frac{1}{n} \sum_{i=1}^n x_{i}, \qquad  %\nonumber \\ 
\sigma_x^2(\mathbf{x}) &= \frac{1}{n} \sum_{i=1}^n (x_{i} - \overline{x})^2,  \nonumber \\ 
 \overline{y}(\mathbf{x}) = \frac{1}{n} \sum_{i=1}^n y_{i}, \qquad  %\nonumber \\ 
\sigma_y^2(\mathbf{x}) &= \frac{1}{n} \sum_{i=1}^n (y_{i} - \overline{y})^2.  
%NOT 1/(N-1) because we are measuring the actual Variance, not the variance of samples drawn from a distribution
\end{align}

@@

If particles with radius $r$ are in a bounded environment with sides of length $[\ell_x, \ell_y]$, the unforced variance asymptotically grows to the variance of a uniform distribution,
\begin{align}
[\sigma_x^2,\sigma_y^2] = \frac{1}{12}[ (\ell_x - 2 r)^2,(\ell_y - 2 r)^2].\label{eq:VarianceUniformDistribution}
\end{align}

 A flat obstacle can be used to decrease variance. Pushing a group of dispersed particles against a flat obstacle will decrease their variance until the minimum-variance (maximum density) packing  is reached. For large $n$, Graham ~\cite{graham1990penny} showed that the minimum-variance packing  for $n$ circles with radius $r$ is 
 \begin{align} \label{eq:optimalvar}
 \sigma^2_{\rm{optimal}}(n,r) \approx   \frac{\sqrt{3}}{\pi} n r^2\approx 0.55 n r^2.
 \end{align} 
% Sloan minimized second moment U = 1/d^2 \sum_{i=1}^{n} || P_i - P ||^2 = (\sqrt{3} n^2)/(4 \pi), where d = 2r.
% Therefore the variance is d^2 (\sqrt{3} n^2)/(4 \pi), or 4^r^2 (\sqrt{3} n^2)/(4 \pi) = r^2 (\sqrt{3} n^2)/(\pi)If we only need to reduce variance in a single axis, the variance can be reduced to zero, given sufficient space.

@@

Similar to the PD controller in Eq.~\eqref{eq:PDcontrolPosition} that controls the mean particle position, a controller to regulate the variance to $\sigma^2_{\rm{ref}}$ is:
\begin{align}
u_x &= K_{p}(x_{\rm{goal}}(\sigma^2_{\rm{ref}}) - \bar{x}) - K_{d}\bar{v}_x + K_{i}(\sigma^2_{\rm{ref}}-\sigma^2_{x}),   \label{eq:PDcontrolVariance}\\
u_y &= K_{p}(y_{\rm{goal}}(\sigma^2_{\rm{ref}})  - \bar{y}) - K_{d}\bar{v}_y + K_{i}(\sigma^2_{\rm{ref}}-\sigma^2_{y}).  \label{eq:PDcontrolVarianceIny}
\end{align}
We call the gain scaling the variance error $K_i$ because the variance, if unregulated, integrates over time. This controller requires a vertical and a horizontal wall.
Eq.~\eqref{eq:PDcontrolVariance} assumes the nearest wall is to the left of the particle at $x=0$, and chooses a reference goal position that in steady-state would have the correct variance according to \eqref{eq:VarianceUniformDistribution}:
\begin{align}
x_{\rm{goal}}(\sigma^2_{\rm{ref}}) = \ell_x/2 = r + \sqrt{3\sigma^2_{\rm{ref}}}.
\end{align}
 If a wall to the right is closer, the signs of $[K_p,K_i]$ are inverted, and the location $x_{\rm{goal}}$ is translated.  A similar argument applies to~\eqref{eq:PDcontrolVarianceIny}.

@@

A key challenge is to select proper values for $\sigma_{\rm{min}}^2$ and $\sigma_{\rm{max}}^2$.  The optimal packing variance was given in \eqref{eq:optimalvar}.
The random packings generated by pushing our particles into corners are suboptimal, so we choose the conservative values: 
%shown in Fig.~\ref{fig:hysteresis}:
\begin{align} \label{eq:sigMaxMin}
 \sigma^2_{\rm{min}} &= 2.5r+ \sigma^2_{\rm{optimal}}(n,r), \nonumber\\
  \sigma^2_{\rm{max}} &= 15r+ \sigma^2_{\rm{optimal}}(n,r).
  \end{align}
%\sigma \approx 0.371258 n r,   \sigma_{min} = n*r*1/2, \sigma_{min} = n*r*3/2, 
%  this result was verified in http://www-math.mit.edu/~tchow/penny.pdf
